Causal Forest Approach for Site-specific Input Management via On-farm Precision Experimentation

### 2022-02-06 Writing/CEA_2nd_round/replies_rev_1.rmd
restructuring of the folders and files

**1. The first point of the highlight describes the advantages of the CF model over other machine learning methods. The advantages of the CF model are not suitable for the highlight of this study. Therefore, I suggest that the authors revise the first point of the highlight.**
<span style="color:blue">

>>>>

**1. The first point of the highlight describes the advantages of the CF model over other machine learning methods. The advantages of the CF model are not suitable for the highlight of this study. Therefore, I suggest that the authors revise the first point of the highlight.**

<span style="color:blue">  
Yes, we agree with your point, and we have decided to simply drop it from our highlights.

### 2022-02-06 Writing/manuscript_cea.rmd
restructuring of the folders and files

```{r echo = F, results='hide', message=FALSE}
library(knitr)
# library(here)

# opts_knit$set(root.dir = here())
opts_knit$set(root.dir = "~/Dropbox/ResearchProject/CF_for_VRA")

@@

```{r analyses, results = "hide", fig.show='hide', cache = FALSE}
#/*=================================================*/
#' # Preparation
#/*=================================================*/
# This is where you prepare data sets and results 
# library(here)
# setwd(here())
#/*----------------------------------*/
#' ## Create figures and tablesb
#/*----------------------------------*/
# --- Convert an Rmd file into a pure R file using purl() ---#
knitr::purl("./GitControlled/Codes/PrepareResults_1.Rmd", output = "./GitControlled/Codes/PrepareResults_1.R")

# --- Read the simulation results and prepare figures and tables ---#
source("./GitControlled/Codes/PrepareResults_1.R")
```

@@

## Estimating EOIR and Treatment Effects

@@

Identity \@ref(eq:eqn2) states that the necessary elements of the information needed for profit maximization are the input price, the output price, and the partial derivative of the yield function with respect to the input. While knowledge of the yield function $f(I, \mathbf{c})$ in Eq. \@ref(eq:eqn1)) is sufficient to solve the maximization problem above (since its derivative can be calculated), identity \@ref(eq:eqn2) shows that it is not a necessary condition; rather, knowing the _change_ in yield in response to a change in input rate is.

@@

CF was developed specifically for identifying heterogeneous treatment effects [@athey2016recursive; @Wager2018a]. Like RF, CF works by building an ensemble of trees, called *causal trees* (CT), by recursively partitioning covariate space into two leaves, where each leaf estimates conditional average treatment effects as the mean differences between the dependent variable (here, yield) of the treated and control observations that belong to the leaf. Figure \@ref(fig:ct-ex) shows an example of CT where the treatment is an increase in seed rate from $27,000$ seeds/ha (control) to $31,000$ seeds/ha (treated) with apparent electrical conductivity ($\textit{ecs}$)($dS/m$) and topographical $\textit{slope}$ ($degree$) as variables to explain the heterogeneity of the impact of the treatment impact. 
In the first node, all the samples ($1024$ observations) belong to the node, and the causal effect of the treatment was calculated as the difference in the mean yield of the treated and control groups, which turned out to be $593$ kg/ha in this example.

```{r ct-ex, fig.width=4, fig.height=4, fig.align="center", fig.cap = "An illustrative example of causal tree"}
# knitr::include_graphics("./GitControlled/Writing/Re_JournalSubmission/ctree_ex.pdf")
readPNG("./GitControlled/Writing/Re_JournalSubmission/ctree_ex.png") %>% grid.raster()
```

The first data split criterion is whether the $\textit{ecs}$ value is greater or less than $25$. The observations with $\textit{ecs}$ greater than or equal to $25$ move to the left node, while those with $\textit{ecs}$ less than $25$ move to the right. For each of the subsamples, the treatment effect is identified as the difference in the mean yields of the treated and control groups within the subsample. The treatment effect for the low-$\textit{ecs}$ observations was $619$ kg/ha, and for the high-$\textit{ecs}$ observations was $982$ kg/ha. This means that for the impact of the increase in seed rate the treatment effect is heterogeneous over space, having a larger positive impact on yield in high-$\textit{ecs}$ areas. Trees are further developed to have more groups of subsamples to allow for more flexible representations of the heterogeneous treatment effects. The estimated heterogeneous treatments effects can then be used to identify site-specific EOIRs. It is noteworth that at no point CF estimates yield _level_.

@@

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate site-specific EONRs. The simulations’ first step generated one thousand synthetic fields that "resembled" an actual field. Then, for each simulation round, one of the fields was used as a training data set, and another field that is different from the one used as a training data set in that round was used as a validation data set. Field characteristics and yield were simulated assuming the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training data set, and checked the statistical performance of the predicted site-specific EONR using the validation data set. Detailed descriptions of these steps are provided below. The *R* and *python* codes that implement the simulation analyses are available at https://github.com/tmieno2/ML_VRA.git.

@@

```{r field-map, out.width="50%", fig.cap = "Partitioning of field into plots, subplots and cells"}
# knitr::include_graphics("./GitControlled/Writing/Re_JournalSubmission/field_devision_v2.png")
readPNG("./GitControlled/Writing/Re_JournalSubmission/field_devision_v2.png") %>% grid.raster()
```

@@

```{r fig-parameters, fig.width=6, fig.height=5, fig.cap = "An Illustrative Example of Spatial Distributions of Field Characteristics"}
# field_alpha + field_beta + field_ymax
grid.arrange(field_ymax, field_alpha, field_beta, field_m_error, ncol=2, nrow=2)
```

@@

### Modeling Scenarios

@@

### Estimation of Site-specific EONRs by RF, BRF, and CNN

The RF, BRF, and CNN methods all followed the same conceptual steps to estimate site-specific EONR: 1) estimate the yield production functions and 2) calculate site-specific EONRs based on that estimated function. For the given modeling scenario, RF and BRF predict yield by using all the available covariates as explanatory variables in the estimation process. The `grf` package (version 1.2.0 [@tibshirani2018package]) in R was used for RF and BRF modeling.

@@

### Estimation of Site-specific EONRs by CF

@@

With the treatment effects from the base obtained, site-specific EONRs were identified by finding the N rate resulting in the highest profit. Yield levels were not predicted at any point in the CF-based approaches.

RF, BRF, and CF from the `grf` package have several tunable hyperparameters (e.g., the minimum node size in each tree, the number of covariates used for node splitting, and parameters involving honest tree-building process), and the optimal values for these parameters were selected by cross-validation. It was also necessary to set the number of trees grown in a forest (*num.trees*). Individual trees in a forest were built with randomly subsampled observations, leading to a different prediction in every forest even when the same data set was used. To reduce the variance of predictions for prediction accuracy, we needed to assign *num.trees* a high value, $4000$.

@@

Virtually all the previous studies have used yield prediction accuracy to rank and select models. This is understandable, since whereas yield can be "ground-truthed" the true causal impact of a treatment cannot be determined from real-world data, making it impossible to cross-validate EONR. 
However, the simulation results reported here show clearly that using a model’s yield prediction accuracy as a model selection criterion is not well justified if the ultimate objective is accurate estimation of site-specific EONRs.

@@

We examined the use of CF-base for site-specific input management recommendations. Using Monte Carlo simulations under various modeling scenarios, we compared the CF estimations of site-specific EONRs to those of the prediction-oriented ML methods RF, BRF, and CNN. CF-base consistently outperformed the other methods across all the modeling scenarios in terms of EONR prediction and profit generation. We also showed that the model’s yield prediction accuracy was uncorrelated with its EONR prediction accuracy, implying that yield prediction accuracy should not be used as a model selection criterion if the ultimate goal of the experiment is to create a site-specific EOIR map.

@@

>>>>

```{r echo = F, include = FALSE}
library(knitr)
library(here)
here::i_am("GitControlled/Writing/manuscript_cea.rmd")
# opts_knit$set(root.dir = here())

@@

```{r analyses, results = "hide", fig.show='hide', cache = FALSE}
# /*=================================================*/
#' # Create figures and tables
# /*=================================================*/
# --- Convert an Rmd file into a pure R file using purl() ---#
knitr::purl(
  here("GitControlled/Codes/PrepareResults.rmd"),
  output = here("GitControlled/Codes/PrepareResults.R")
)

# --- Read the simulation results and prepare figures and tables ---#
source(here("GitControlled/Codes/PrepareResults.R"))
```

@@

## EOIR Estimation and Treatment Effects

@@

Identity \@ref(eq:eqn2) states that the necessary elements of the information needed for profit maximization are the input price, the output price, and the partial derivative of the yield function with respect to the input. While knowledge of the yield function $f(I, \mathbf{c})$ in Eq. \@ref(eq:eqn1)) is sufficient to solve the maximization problem above (since its derivative can be calculated), identity \@ref(eq:eqn2) shows that it is not a necessary condition; rather, knowing the "change" in yield in response to a change in input rate is.

@@

CF was developed specifically for identifying heterogeneous treatment effects [@athey2016recursive; @Wager2018a]. Like RF, CF works by building an ensemble of trees, called *causal trees* (CT), by recursively partitioning covariate space into two leaves, where each leaf estimates conditional average treatment effects as the mean differences between the dependent variable (here, yield) of the treated and control observations that belong to the leaf (please see Appendix for an example of a causal tree).

@@

Monte Carlo simulations were conducted for nitrogen experiments for corn production to estimate site-specific EONRs. The simulations' first step generated one thousand synthetic fields that "resembled" an actual field. Then, for each simulation round, one of the fields was used as a training data set, and another field that is different from the one used as a training data set in that round was used as a validation data set. Field characteristics and yield were simulated assuming the Mitscherlich-Baule (MB) yield response function. We used the RF, BRF, and CNN ML methods to estimate yield response functions, and we used the CF method to estimate treatment effects. For each method, we estimated site-specific EONRs using the training data set, and checked the statistical performance of the predicted site-specific EONR using the validation data set. Detailed descriptions of these steps are provided below. The *R* and *python* codes that implement the simulation analyses are available at https://github.com/tmieno2/ML_VRA.git.

@@

```{r field-map, out.width="50%", fig.cap = "Partitioning of field into plots, subplots and cells"}
# knitr::include_graphics("./GitControlled/Writing/field_structure.png")
readPNG(here("GitControlled/Writing/field_structure.png")) %>% grid.raster()
```

@@

```{r fig-parameters, fig.width=6, fig.height=5, fig.cap = "An Illustrative Example of Spatial Distributions of Field Characteristics", cache = TRUE}
# field_alpha + field_beta + field_ymax
grid.arrange(field_ymax, field_alpha, field_beta, field_m_error, ncol = 2, nrow = 2)
```

@@

## Modeling Scenarios

@@

### Training and Estimation of Site-specific EONRs by RF, BRF, and CNN

The RF, BRF, and CNN methods all followed the same conceptual steps to estimate site-specific EONR: 1) estimate the yield production functions and 2) calculate site-specific EONRs based on that estimated function. For the given modeling scenario, RF and BRF predict yield by using all the available covariates as explanatory variables in the estimation process. The `grf` package (version 1.2.0 [@tibshirani2018package]) in R was used for RF and BRF modeling. RF, BRF, and CF from the `grf` package have several tunable hyperparameters (e.g., the minimum node size in each tree, the number of covariates used for node splitting, and parameters involving honest tree-building process), and the optimal values for these parameters were selected by cross-validation. Individual trees in a forest were built with randomly subsampled observations, leading to a different prediction in every forest even when the same data set was used. To reduce the variance of predictions for prediction accuracy, we set the number of trees (*num.trees*) at a high value, $4000$.

@@

### Training and Estimation of Site-specific EONRs by CF

@@

With the treatment effects from the base obtained, site-specific EONRs were identified by finding the N rate resulting in the highest profit for each $i$. Yield levels were not predicted at any point in the CF-based approaches.

@@

Virtually all the previous studies have used yield prediction accuracy to rank and select models. This is understandable, since whereas yield can be "ground-truthed" the true causal impact of a treatment cannot be determined from real-world data, making it impossible to cross-validate EONR. 
However, the simulation results reported here show clearly that using a model's yield prediction accuracy as a model selection criterion is not well justified if the ultimate objective is accurate estimation of site-specific EONRs.

@@

We examined the use of CF-base for site-specific input management recommendations. Using Monte Carlo simulations under various modeling scenarios, we compared the CF estimations of site-specific EONRs to those of the prediction-oriented ML methods RF, BRF, and CNN. CF-base consistently outperformed the other methods across all the modeling scenarios in terms of EONR prediction and profit generation. We also showed that the model's yield prediction accuracy was uncorrelated with its EONR prediction accuracy, implying that yield prediction accuracy should not be used as a model selection criterion if the ultimate goal of the experiment is to create a site-specific EOIR map.

@@

# (APPENDIX) Appendix {-}

# Appendix A: Causal Tree Example {-}

Figure \@ref(fig:ct-ex) shows an example of CT where the treatment is an increase in seed rate from $27,000$ seeds/ha (control) to $31,000$ seeds/ha (treated) with apparent electrical conductivity ($\textit{ecs}$)($dS/m$) and topographical $\textit{slope}$ ($degree$) as variables to explain the heterogeneity of the impact of the treatment impact. In the first node, all the samples ($1024$ observations) belong to the node, and the causal effect of the treatment was calculated as the difference in the mean yield of the treated and control groups, which turned out to be $593$ kg/ha in this example.

```{r ct-ex, fig.width=4, fig.height=4, fig.align="center", fig.cap = "An illustrative example of causal tree"}
# knitr::include_graphics("./GitControlled/Writing/ctree_ex.pdf")
readPNG(here("GitControlled/Writing/ctree_ex.png")) %>% grid.raster()
```

The first data split criterion is whether the $\textit{ecs}$ value is greater or less than $25$. The observations with $\textit{ecs}$ greater than or equal to $25$ move to the left node, while those with $\textit{ecs}$ less than $25$ move to the right. For each of the subsamples, the treatment effect is identified as the difference in the mean yields of the treated and control groups within the subsample. The treatment effect for the low-$\textit{ecs}$ observations was $619$ kg/ha, and for the high-$\textit{ecs}$ observations was $982$ kg/ha. This means that for the impact of the increase in seed rate the treatment effect is heterogeneous over space, having a larger positive impact on yield in high-$\textit{ecs}$ areas. Trees are further developed to have more groups of subsamples to allow for more flexible representations of the heterogeneous treatment effects. The estimated heterogeneous treatments effects can then be used to identify site-specific EOIRs. It is noteworth that at no point CF estimates yield "level".
